---
title: "NTS2Source"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}

library(DiagrammeR)

```


# Concept of NTS2Source

```{r}

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = true]
      
# node statements
node [shape = box,
fontname = Helvetica, fontsize = 20]

# Sample information
SAMP1 [label = 'Sample_Id']
SAMP2 [label = 'Sample_MetaData']

# Data pretreatment and data analysis
DATA1 [label = 'MS_Ready']

# Chemical analysis
ANAL1 [label = 'Analysis_MassSpec']
ANAL2 [label = 'Quantification']

# Chemical information
CHEM1 [label = 'Chemical_Id']
CHEM2 [label = 'Chemical_PhysicalProp']

# Various sources of chemicals


# Libraries should be from different sources: NIST, Mona, MassbankEU, InHouse,..
LIB1 [label = 'Library_Spec']
LIB2 [label = 'Library_Toxcast']

# List such as NormaNews suspect list, InHouse,..
LIST1 [label = 'List_Susp']
LIST2 [label = 'List_TransfProd_Pred']
LIST3 [label = 'Source_Products']

# Different output results, such as msp of spectrum
RES1 [label = 'Output_RawData']

# Various algorithm to match similarities, cosine similarity,..
MOD1 [label = 'Algorithm_Match']
MOD2 [label = 'Nontarget_HCA']
MOD3 [label = 'Nontarget_MassDefect']

ID1 [label = 'Identification']

EXPO1 [label = 'Exposure_IndoorDust']


###### edge statements ######

CHEM1 -> CHEM2
SAMP2 -> SAMP1
SAMP1 -> ANAL1
ANAL1 -> RES1 [label = 'GIVES']


LIST1 -> DATA1
{LIB1 DATA1} -> MOD1
RES1 -> {MOD1 MOD2 MOD3}
{MOD1 MOD2 MOD3} -> ID1 [label = 'GIVES']

ID1 -> CHEM1 [label = 'GIVES']
CHEM1 -> LIST3 [label = 'FOUND_IN']
CHEM1 -> LIB2

CHEM1 -> LIST2 -> RES1
{CHEM1 ID1} -> ANAL2

{ANAL2 CHEM2 LIB2} -> EXPO1
      
      }
    
")

 



```




# Ontology  
https://www.ebi.ac.uk/ols/ontologies/cheminf  
http://www.obofoundry.org/ontology/cheminf.html  


# Edge labels:  
- GIVES:  
- FOUND_IN:  


Create graph database using Neo4j desktop

Copy the files to the local import folder (in this case it was at "D:\Program\Neo4js_data\relate-data\dbmss\dbms-a42acde1-a3d0-4c0c-930f-99319d3b5998")

LOAD csv file for Sundahus database:  

```{r eval=FALSE}

LOAD CSV FROM "file:///SH_data.csv" AS line FIELDTERMINATOR '\u0009' CREATE (:Products
  {Varumarke: line[1],
  Namn: line[2],
  Beskrivning: line [3],
  Artikelnummer: line [4],
  BSAB_kod: line [5],
  BK04_1: line [6],
  BK04_2: line [7],
  BK04_3: line [8]})
  
```

DELETE node
```{r eval=FALSE}

MATCH (n: Products) DELETE (n)

```


CREATE node

```{r eval=FALSE}

CREATE (:Chemical 
      {Chemical_name: "BDE-99",
       CAS: "3434-343-33"})

```

CREATE links between nodes

```{r eval=FALSE}

MATCH (a:Chemical), (b:Products)
WHERE a.Chemical_name = "BDE-99" AND b.Namn = "Vattenburen klarlack for brandisolering av tra, plywood, spanskivor och andra traprodukter utomhus."
CREATE (a)-[r:RELTYPE {name: a.name + '<->' + b.name}]->(b)
RETURN type(r), r.nam

```



```{r eval=FALSE}

CREATE p = (Chemical {Chemical_name:"BDE-99"})-[:FOUND_IN]->(Products)
```






# Background on graph database

Knowledge Graphs are effective tools to organize data in an unbiased and unsupervised manner. The 
main difference between conventional databases and Knowledge Graph is that in a conventional 
database one needs to know the question in order to query the database and retrieve the correct 
answer. Knowledge Graphs can be analysed to identify unknown patterns of relatedness between the 
data points, which could not be extrapolated when considering them individually. The computational 
AI-based framework needed to learn from Knowledge is well established and successfully used in various fields, but thus far not fully exploited in toxicology.
Watson Health https://www.ibm.com/watson-health ; Google Knowledge Graph https://www.google.com ; IBM RXN https://rxn.res.ibm.com/ ;
https://www.telegraph.co.uk/technology/2021/01/18/drugs-will-designed-ai-decades-end/#comment

The nodes of the RESONATE Knowledge Graph will be chemicals, exposures, doses, genes, proteins, cells, tissues,
diseases, regulatory requirements, while the connections between the nodes will be different aspects of their
relationships. For instance, chemicals will share edges with genes whose expression is altered by the exposure, as
recorded in manually curated toxicogenomic data. Likewise, tissues, cell lines or in vitro models will share
connections with genes (based on the expression patterns recorded in relevant omics datasets). Based on the set of
edges explicitly present in the Knowledge Graph (e.g., chemical.X - gene.A - cell.Y - phenotype.W), we will first
extrapolate the intrinsic edges that allow a simplified representation of the data (eg. chemical.X - cell.Y; chemical.X
- phenotype.W; etc). Next, by the use of specific AI algorithms that are able to learn from the structure of the graph,
we will infer new edges through link prediction, leading to further new discoveries. The first and foremost question
we want to answer with this approach is: “is there an interaction between exposure X and Y?”, and next: “is the effect
of the interaction additive, antagonistic or synergistic?” This includes analysis of the quality of the interaction
between components of a mixture, predicted/extracted from data generated by the exposure, hazard, risk and IATA
modelling in WPs2-4. Additional relevant aspects that we will be able to address by learning from the RESONATE
Knowledge Graph are, for instance, related to which in vitro system is the best to investigate a phenotype of interest,
which characteristics need a computational model to be regulatory-ready, or which exposures of mixture of exposures
are more probable to cause a certain phenotype in individuals with a certain disease. By continuously updating the
Knowledge Graph with outputs from the predictive AI models, the ability of the Knowledge Graph to support
interpretation of mixture effects will rapidly increase.
